---
title: "Problem Set 2"
author: "Leah Shiferaw"
date: "2024-01-31"
output: word_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(tidyverse)
library(knitr)

df = read_csv('greenspace_data_share.csv')
 
```

Answer the questions below to the best of your ability. Write clearly, and format your tables and visuals appropriately. You must use `R Markdown` to compose and compile your work. For full credit, `echo` all code chunks, and include your `setup` chunk. Submit your work in hard copy at the beginning of class. 

You need the [Global Greenspace Indicator Data](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/TMWYHB) for this assignment. Review the `README.txt` file for information about the data, variables, etc.

1. Show me that you're all set on GitHub. Create a public repository named `PSet2`. Clone it, include all your project files (including your `.Rmd` and `.pdf` files) for the work below, and commit/push your work to your repository. Include the link to your repo as your answer to this question. 

https://github.com/leah-shiferaw/PSet2.git


2. The script below doesn't work. Type the corrected code chunk into your problem set. Annotate any line you correct to note your fix (i.e. `# unbalanced parentheses`). *There are more than five errors.*

```{r}

  library(tidyverse)

# open my data
  gspace = read_csv('greenspace_data_share.csv')
  
#added quotation marks to greenspace_data_share.csv
  
# summarize average urban greenspace by region
  table = 
    gspace |>
    group_by(Major_Geo_Region) |>
    summarise(
      obs = n (),
      avg = mean(annual_avg_2020),
      weightedavg = mean(annual_weight_avg_2020)
    ) 
  

#added comma after observation, underscore to annual_avg_2020 and pipe operator after Major_Geo_Region, and removed space between weighted average

```

3. How many urban areas does the greenspace data cover?


  The data covers 1038 urban areas.


4. In a couple of sentences and with reference to a well-formatted tabulation, describe the greenspace classification scores for urban areas in 2021.

```{r}

table2 <- gspace |>
  group_by(Major_Geo_Region) |>
  summarise(
    obs = n(),
    avg = mean(annual_avg_2021),
    weightedavg = mean(annual_weight_avg_2021)
  )

```

  Urban areas in Latin America, Europe, and Asia had a lower weighted average of greenspaces. Urban areas in Oceania had the highest weighted average of greenspaces. 
  
5. Report the number of urban areas that satisfy the conditions below. Either write your code inline or echo the code that generated the answer.   

    
    a. Scored `High` or above for greenspace in 2015. 
  
        table3 <- gspace |>
        count(indicator_2015)
 
        66 Urban areas scored "high" or "very high"
    
     b. Scored `Exceptionally Low` at any point in the years covered. 

  
  filter(gspace, indicator_2010 == "Exceptionally Low" |  
  indicator_2015 == "Exceptionally Low" |             
  indicator_2020 == "Exceptionally Low" | 
  indicator_2021 == "Exceptionally Low") |>
    summarise(
      obs = n())
      
   240 areas scored "exceptionally low" from 2010 to 2021.
   
    c. Urban areas in arid climate that became greener (as measured by annual weighted average) from 2010 to 2020. 


filter(gspace, Climate_region == "Arid",) |>
    summarise(
      obs = n()
      )
  
  225 urban areas became greener from 2010 to 2020.
  
6. How many urban areas became less green (measured by annual average) from 2010 to 2021? Were these changes concentrated in a particular geographic or climate region? Explain (with evidence, of course).

```{r}
gspace <- transform(gspace, 
    annualaverage = annual_avg_2021 - annual_avg_2010
    )
    
table4= 
    gspace |>
    group_by(Major_Geo_Region) |>
    filter(annualaverage < 0 ) |> 
    summarise( 
      obs = n()
      )

  table5= 
    gspace |>
    group_by(Climate_region) |>
    filter(annualaverage < 0 ) |> 
    summarise( 
      obs = n()
      )

```

 128 areas became less green over time, and this occurred in temperate and continental regions. 

  
7. Present a histogram showing the change in greenspace (annual average) from 2010 to 2021. Note that you will need to create a new variable equal to this difference.  

```{r}
hist(
  gspace$annualaverage,
  col = 'blue',
  main = 'Change in Greenspace 2010 to 2021',
  xlab = 'Annual Average',
  ylab = 'Peak Weight',
  border = 'black'
)

```

8. Present a scatter plot of population weighted greenspace in 2021 over the greenspace in 2010. 

```{r}

ggplot(df, aes(x = annual_weight_avg_2010, y = annual_weight_avg_2021)) +
geom_point () +
labs(
    x = 'Annual Weight Average 2010',
    y = 'Annual Weight Average 2021'
  )

```


**BONUS OPPORTUNITY**: Use color-coding to differentiate urban areas that added versus lost greenspace in that time. Then include a 45-degree line to further highlight the change.


```{r}

ggplot(df, aes(x = annual_weight_avg_2010, y = peak_weight_2021)) +
geom_point (color = "red") +
geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "orange") +
labs(
    x = 'Annual Weight Average', 
    y = 'Peak Weight 2021' , 
  )

```



